Comparison of dimensionality reduction schemes for parallel global optimization algorithms \thanks{The study was supported by the Russian Science Foundation, project No 16-11- 10150}

### 2018-05-15 2018/RSD/MultiEvolvents/RSD.tex
RSD: fix review issues

Such set of evolvents can be produced by shifting the source evolvent $y^0(x)$ by $2^{-l},0
\leq l \leq L$ on each coordinate. Each evolvent has it's own corresponding hypercude $D_l=
\left\{y \in R^N: -2^{-1} \leq y_i+2^{-l} \leq 3 \cdot 2^{-1},\ 1\leq i\leq N\right\},\ 0 \leq l \leq
L$.

@@

As it has been already mentioned above (Sec.~\ref{sec:shifted}), the loss of information on the
proximity of the points in the multidimensional space could be compensated in part by the use
of multiple mappings $Y_L(x)=\{y^1(x),...,y^L(x)\}$. However, the Peano-type curve preserves
a part of this information itself: it is not an injective mapping. Therefore, if a single image
$y(x)\in \mathbb{R}^N$ is available, on can obtain several different preimages
$t_j\in[0,1], t_j \not = x$, which could be added into the search information of the method later.

@@

%------------------------------------------------------------------------------
\subsection{Parallel algorithm exploiting a set of evolvents}
\label{sec:parallel_evolvents}
Using the multiple mapping allows solving initial problem (\ref{eq:task}) by parallel solving the
problems
\[
\min\{\varphi(y^s(x)):x\in [0,1]\}, 1\leqslant s\leqslant S
\]
on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a
separate processor. The trial results at the point \(x^k\) obtained for the problem being solved by
particular processor are interpreted as the results of the trials in the rest problems (in the
corresponding points \(x^(k_1 ),\dots,x^(k_S))\). In this approach, a trial at the point \(x^k \in
[0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of
operations.
\par
1. Determine the image \(y^k=y^s (x^k)\) for the evolvent \(y^s (x)\).
\par
2. Inform the rest of processors about the start of the trial execution at the point\( y^k\) (the
blocking of the point \(y^k\) ).
\par
3. Determine the preimages \(x{}^{k_s}  \in [0,1], 1\leqslant s\leqslant S\), of the point \(y^k\) and interpret the
trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points
\(x{}^{k_1} ,\dots,x{}^{k_s} \)
\par
4. Inform the rest of processors about the trial results at the point \(y^k\).
\par

@@

The comparison of the global optimization algorithms was performed by the evaluation of the
quality of solving a set of problems from some test class.
In the present paper, the test class generated by GKLS generator~\cite{Gaviano2003} was
considered. This generator allows constructing the complex multiextremal problems of various
dimensions. In the present work, the series of 100 problems from the classes of the dimensions
of 2, 3, 4, and 5 were considered.
Each class had two degrees of complexity --- \textit{Simple} and \textit{Hard}. The parameters
of the generator for the considered classes were given in Ref.~\cite{Gaviano2003}.

>>>>

Such set of evolvents can be produced by shifting the source evolvent $y^0(x)$ by $2^{-l},0
\leq l \leq L$ on each coordinate. Each evolvent has it's own corresponding hypercube $D_l=
\left\{y \in R^N: -2^{-1} \leq y_i+2^{-l} \leq 3 \cdot 2^{-1},\ 1\leq i\leq N\right\},\ 0 \leq l \leq
L$.

@@

As it has been already mentioned above (Sec.~\ref{sec:shifted}), the loss of information on the
proximity of the points in the multidimensional space could be compensated in part by the use
of multiple mappings $Y_L(x)=\{y^1(x),...,y^L(x)\}$. However, the Peano-type curve preserves
a part of this information itself: it is not an injective mapping. Therefore, if a single image
$y(x)\in \mathbb{R}^N$ is available, one can obtain several different preimages
$t_j\in[0,1], t_j \not = x$, which could be added into the search information of the method later.

@@

%------------------------------------------------------------------------------
\subsection{Parallel algorithm exploiting a set of evolvents}
\label{sec:parallel_evolvents}
Using the multiple mapping allows solving initial problem (\ref{eq:task}) by parallel solving the
problems
\[
\min\{\varphi(y^s(x)):x\in [0,1]\}, 1\leqslant s\leqslant S
\]
on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a
separate processor. The trial results at the point \(x^k\) obtained for the problem being solved by
particular processor are interpreted as the results of the trials in the rest problems (in the
corresponding points \(x^{k_1},\dots,x^{k_S})\). In this approach, a trial at the point \(x^k \in
[0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of
operations.
\par
1. Determine the image \(y^k=y^s (x^k)\) for the evolvent \(y^s (x)\).
\par
2. Inform the rest of processors about the start of the trial execution at the point \( y^k\) (the
blocking of the point \(y^k\) ).
\par
3. Determine the preimages \(x{}^{k_s}  \in [0,1], 1\leqslant s\leqslant S\), of the point \(y^k\) and interpret the
trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points
\(x{}^{k_1} ,\dots,x{}^{k_s} \)
\par
4. Inform the rest of processors about the trial results at the point \(y^k\).
\par

@@

The comparison of the global optimization algorithms was performed by the evaluation of the
quality of solving a set of problems from some test class.
In the present paper, the test class generated by GKLS (Gaviano, Kvasov, Lera, Sergeyev) generator~\cite{Gaviano2003} was
considered. The generator creates objective functions by distorting a convex quadratic function by polynomials in order to introduce local minima.
Thus, GKLS allows constructing the complex multiextremal problems of various dimensions.
In the present work, the series of 100 problems from the classes of the dimensions
of 2, 3, 4, and 5 were considered.
Each class had two degrees of complexity --- \textit{Simple} and \textit{Hard}. These the classes have different
radius of the attraction region of the global minimizer (\textit{Hard} has smaller region) and distance from the global minimizer to the vertex of the quadratic function (\textit{Simple} has smaller distance).
The parameters of the generator for the considered classes were given in Ref.~\cite{Gaviano2003}.
